# Multilevel modeling of ego-network data {#multilevel}

```{r include=FALSE, cache=FALSE}
knitr::read_chunk("./Scripts/06_Multilevel.R")
```

In progress.

## Resources

### Multilevel models for egocentric network data

Specific resources on multilevel modeling of egocentric/personal network data:

* [@vacca_multilevel_2018](https://www.sa-ijas.org/ojs/index.php/sa-ijas/article/view/30-3): overview of multilevel modeling for egocentric/personal network data, with R demo (the code below is drawn in part from this article).
* @simonoff_multilevel_2013: overview of multilevel modeling for egocentric and sociocentric data.
* [@vacca_cross-classified_2019](http://journals.sagepub.com/doi/10.1177/0049124119882450): multilevel models for overlapping egocentric networks.
* @mccarty_conducting_2019: Ch. 13.
* @perry_egocentric_2018: Ch. 8.
* @crossley_social_2015: Ch. 6.

### Multilevel modeling in general

General resources on multilevel models:

* [@rasbash_lemma:_2008](https://www.cmm.bris.ac.uk/lemma): Extensive online course on multilevel modeling for the social sciences, including Stata and R implementation.
* @snijders_multilevel_2012: mode in-depth statistical treatment, social science focus.
* @goldstein_multilevel_2010
* @fox_r_2018: Ch. 7, R implementation with `lme4`.
* [GLMM FAQ](http://bbolker.github.io/mixedmodels-misc/glmmFAQ.html) by Ben Bolker and others: online list of R packages and resources on linear and generalized linear mixed models.

## Prepare the data

```{r mm-setup, include=TRUE, cache=FALSE, tidy= FALSE, message=FALSE}
```

## Random intercept models

```{r rand-intercept, include=TRUE, cache=FALSE, tidy= FALSE}
```

## Random slope models

```{r rand-slope, include=TRUE, cache=FALSE, tidy= FALSE}
```

## Tests of significance

```{r test-sign, include=TRUE, cache=FALSE, tidy= FALSE}
```