# Multilevel modeling of ego-network data {#multilevel}

```{r include=FALSE, cache=FALSE}
knitr::read_chunk("./Scripts/06_Multilevel.R")
```

## Overview

This chapter is still in progress, in the meantime see the multilevel modeling slides in the [workshop folder](http://www.raffaelevacca.com/teaching/egocentric-r/folder/) (`Slides` subfolder). @vacca_multilevel_2018 provides a review and illustration of multilevel modeling for ego-network data. The code below is drawn in part from that article.

## Prepare the data

```{r mm-setup, include=TRUE, cache=FALSE, tidy= FALSE, message=FALSE}
```

## Random intercept models

```{r rand-intercept, include=TRUE, cache=FALSE, tidy= FALSE}
```

## Random slope models

```{r rand-slope, include=TRUE, cache=FALSE, tidy= FALSE}
```

## Tests of significance

```{r test-sign, include=TRUE, cache=FALSE, tidy= FALSE}
```

## Further readings

In order of relevance to this chapter:

+ @vacca_multilevel_2018
+ @mccarty_conducting_2019, Ch. 13.
+ @perry_egocentric_2018, Ch. 8.
+ @crossley_social_2015, Ch. 6.
+ @vacca_cross-classified_2019
+ @simonoff_multilevel_2013 
+ @rasbash_lemma:_2008
+ @snijders_multilevel_2012